A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2024-04-23 tidyindex.qmd
add references and connect sentences in intro

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This violates *tidy principles* [@tidyverse] which makes it difficult to investigate the nature of calculations through a principled data pipeline. It also violates statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric [@???Spiegelhalter, someone else].

>>>>

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This echoes an issue highlighted by @donoho_50_2017, where varying data analysis workflows can resulted in different, or even reversed, conclusions when the identical data is analyzed in medicine research. In Greater Data Science (GDS) 6: Science About Data Science, Donoho envisions future research on structuring unified workflow to address methodological variation across studies in data science. It also violates statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric [@efron_bootstrap_1979].

### 2024-04-23 tidyindex.tex
add references and connect sentences in intro

The problem is that every index is developed in its own unique way, by
different researchers or organizations, and often indexes designed for
the same purpose cannot easily be compared. This violates \emph{tidy
principles} (Wickham et al. 2019a) which makes it difficult to
investigate the nature of calculations through a principled data
pipeline. It also violates statistical principles, where quantifying and
understanding uncertainty are essential to deciding on a best measure or
metric {[}@???Spiegelhalter, someone else{]}.

@@

Examples of using a pipeline approach for data analysis can be traced
back to the interactive graphics literature, including A. Buja et al.
(1988); Sutherland et al. (2000); Wickham et al. (2009); Xie, Hofmann,
and Cheng (2014). Wickham et al. (2009) argue that whether made explicit
or not, a pipeline has to be presented in every graphics program, and
making them explicit is beneficial for understanding the implementation
and comparing between different graphic systems. While this comment is
made in the context of interactive graphics programs, it is also
applicable generally to any data analysis workflow. More recently, the
tidyverse suite (Wickham et al. 2019b) takes the pipeline approach for
general-purpose data wrangling and has gained popularity within the R
community. The pipeline-style code can be directly read as a series of
operations applied successively on tidy data objects, offering a method
to document the data wrangling process with all the computational
details for reproducibility.

@@

@@

\bibitem[\citeproctext]{ref-tidyverse}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019a.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-wickham_welcome_2019}
---------, et al. 2019b. {``Welcome to the {Tidyverse}.''} \emph{Journal
of Open Source Software} 4 (43): 1686.
\url{https://doi.org/10.21105/joss.01686}.

>>>>

The problem is that every index is developed in its own unique way, by
different researchers or organizations, and often indexes designed for
the same purpose cannot easily be compared. This echoes an issue
highlighted by Donoho (2017), where varying data analysis workflows can
resulted in different, or even reversed, conclusions when the identical
data is analyzed in medicine research. In Greater Data Science (GDS) 6:
Science About Data Science, Donoho envisions future research on
structuring unified workflow to address methodological variation across
studies in data science. It also violates statistical principles, where
quantifying and understanding uncertainty are essential to deciding on a
best measure or metric (Efron 1979).

@@

Examples of using a pipeline approach for data analysis can be traced
back to the interactive graphics literature, including A. Buja et al.
(1988); Sutherland et al. (2000); Wickham et al. (2009); Xie, Hofmann,
and Cheng (2014). Wickham et al. (2009) argue that whether made explicit
or not, a pipeline has to be presented in every graphics program, and
making them explicit is beneficial for understanding the implementation
and comparing between different graphic systems. While this comment is
made in the context of interactive graphics programs, it is also
applicable generally to any data analysis workflow. More recently, the
tidyverse suite (Wickham et al. 2019) takes the pipeline approach for
general-purpose data wrangling and has gained popularity within the R
community. The pipeline-style code can be directly read as a series of
operations applied successively on tidy data objects, offering a method
to document the data wrangling process with all the computational
details for reproducibility.

@@

\bibitem[\citeproctext]{ref-donoho_50_2017}
Donoho, David. 2017. {``50 {Years} of {Data} {Science}.''} \emph{Journal
of Computational and Graphical Statistics} 26 (4): 745--66.
\url{https://doi.org/10.1080/10618600.2017.1384734}.

\bibitem[\citeproctext]{ref-efron_bootstrap_1979}
Efron, B. 1979. {``Bootstrap {Methods}: {Another} {Look} at the
{Jackknife}.''} \emph{The Annals of Statistics} 7 (1): 1--26.
\url{https://doi.org/10.1214/aos/1176344552}.

@@

\bibitem[\citeproctext]{ref-wickham_welcome_2019}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {Tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.
